### 2019-03-07 manuscript/discussion.tex

% - value of approach
The present work shows the importance of considering the trait linked diversification patterns under a multivariate approach.
Species are created and go extinct based un multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms.
In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.

@@

>>>>

% - value of approach
The present work shows the importance of considering the trait linked diversification patterns under a multivariate approach.
Species are created and go extinct based on multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms.
In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.

@@

% Will: I really like the pathways analyses and have a couple comments
%
% 1) Were the "without diversification" and "with diversification" analyses computed
%    using the same MAP transition rate estimates? Transition rates among states will
%    be very different if estimated using a model that is diversification-independent 
%    compared to when estimated with a trait-dependent diversification model. Since the point of 
%    showing these side-by-side seems to be emphasizing the importance of considering trait-dependent
%    diversification, I wonder how different the "without diversification" contributions would be
%    if calculated from transition rates estimated using a diversification-independent model.
%
% 2) I think stochastic mapping would be another useful perspective on the pathways question
%    (and yeah, its a bit late in the game to propose analyses, so maybe next time!). With stochastic 
%    mapping we could use the entire posterior estimate of the transition rates rather than a point
%    estimate. But more importantly we'd get the relative contributions of the pathways estimated 
%    over the actual Solanceae tree (which has a lot of short interdependent branches rather than 
%    hypothetical long branches). 
%
% 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
%    the rest of the discussion, so modify or discard as needed:
Our work shows that a diploid Solanaceae lineage is much more likely to take a
one-step ID/CP pathway to polyploidy rather than a two-step ID/CD/CP pathway. 
We modeled both of these pathways as anagenetic character changes occurring within a species.
However \citet{goldberg_2012} showed that I/C transitions are often associated with speciation events, and similarly \citet{freyman_2017} demonstrated that D/P transitions may also be associated with speciation events. 
Models that fail to consider transitions that occur at speciation (cladogenetic changes) may
make misleading estimates of anagenetic transition rates.
The HiSSE-based models introduced in our work here could be extended to incorporate cladogenetic transitions. 
It is possible that allowing for cladogenetic transitions may enable the one-step pathway
to even further dominate over the two-step pathway to polyploidy, but this remains to be tested in future work.

### 2019-03-07 manuscript/introduction.tex

%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
Species accumulate within the tree of life at different rates.
A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
Dramatic increases in available phylogenetic and phenotypic data %citation
as well as methodological advances %citation
have greatly accelerated the search for such traits that influence diversification traits. Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) . %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context  in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}. Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.

@@

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution. % refs: stebbins, igic
% Gordon comments here: Really? I find this hard to believe.There are a number of transitions among general selfing and outcrossing mating systems, but I think of SI as only occurring in a limiteed number of angiosperm families. Assigned to Boris
Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}. Faster net diversification of self-incompatible state was also found in the family  \textit{Onagraceae} \citep{freyman_2017} , where authors assumed the possibility of hidden states potentially driving diversification patterns. However, studying breeding system with other correlated traits and unobserved states driving diversification patterns simultaneously has not yet being explored. % Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems. I don't think the de Vos paper is dealing with SI systems. If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying. For example, there are a lot of papers trying to link dioecy/hermaphroditism with diversification. (Vamosi papers on dioecy and diversification)
% R- Boris what do you think is the best way to proceed here? Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae

>>>>

%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
Species accumulate within the tree of life at different rates.
A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
Dramatic increases in available phylogenetic and phenotypic data %citation
as well as methodological advances %citation
have greatly accelerated the search for such traits that influence diversification traits.
Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}). %B: the following is now orphaned, because its inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.

@@

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution. % refs: stebbins, igic
% Gordon comments here: Really? I find this hard to believe.There are a number of transitions among general selfing and outcrossing mating systems, but I think of SI as only occurring in a limiteed number of angiosperm families. Assigned to Boris
Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that hidden states could drive diversification patterns \citep{freyman_2018}.
However, studying breeding system with other correlated traits and unobserved states driving diversification patterns simultaneously has not yet being explored. % Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems. I don't think the de Vos paper is dealing with SI systems. If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying. For example, there are a lot of papers trying to link dioecy/hermaphroditism with diversification. (Vamosi papers on dioecy and diversification)
% R- Boris what do you think is the best way to proceed here? Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae

### 2019-03-07 manuscript/main.tex

\noindent
Rosana Zenil-Ferguson$^{1,\dagger}$,
%
\noindent
J.~Gordon Burleigh$^{2}$,
%
\noindent
William A.~Freyman$^{1}$,
%
\noindent
Boris Igi\'c$^{3}$,
%
\noindent
Itay Mayrose$^{4}$,
%
and
Emma E.~Goldberg$^{1}$
%
\end{center}

%
\noindent$^{1}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
\noindent$^{3}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
\noindent$^{4}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
\noindent$^\dagger$\textit{Author for correspondence}.\\
\vfill
\noindent\textit{Running head:} Ploidy and Breeding Systems in Solanaceae
%\vfill

>>>>

\noindent 
Rosana Zenil-Ferguson$^{1,\dagger}$,
%
J.~Gordon Burleigh$^{2}$,
%
William A.~Freyman$^{3}$,
%
Boris Igi\'c$^{4}$,
%
Itay Mayrose$^{5}$,
%
and Emma E.~Goldberg$^{3}$
%
\end{center}
%
\noindent$^{2}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
%
\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
%
\noindent$^{3}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
%
\noindent$^{4}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
%
\noindent$^{5}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
%
\noindent$^\dagger$\textit{Author for correspondence:}.\\

\noindent\textit{Running head:} 
Ploidy and Breeding Systems in Solanaceae
%\vfill
